CUBE: A Memory-lite Cosmological $N$-body Algorithm

### 2017-10-09 ms_caf/ms_caf.tex
add TianNu and TianZero configs in Table 2.

As a $n_\chi=1$, 1D ($d=1$), 4-coarse-cell ($N_c=4$) example, if $$\chi_1=(-128,127,0,60),$$ particle number density $${\rho_c^{\rm 1D}}=(1,0,2,1),$$ then in unit of coarse cells, $$x_1=(0.001953125, 2.998046875, 2.501953125, 3.736328125).$$

@@

We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass. These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE}. Using different number of tiles per image gives exact same results.

\begin{table}[]
\centering
\caption{Simulation configurations}
\label{t.sim}
\begin{tabular}{lrrrrrr}
\hline
& \multicolumn{6}{c}{Configurations}\\
\cline{2-7}
Name  & $N_{\rm node}$ & $L/({\rm Mpc}/h)$ & $z_i$ & $z_{\rm ckpt}$ & $N_p$ & $m_p/\Msun$ \\
\hline
S512   & 8     & 200   & 49 & 0    & $512^3$  & $7.5\times 10^9$    \\
S256   & 1     & 80    & 49 & 0    & $256^3$  & $3.8\times 10^9$    \\
S2048S & 64    & 400   & 49 & 1,0  & $2048^3$ & $9.4\times 10^8$    \\
S2048L & 64    & 1200  & 49 & 1,0  & $2048^3$ & $2.5\times 10^{10}$ \\
\hline
TianNu & 13824 & 1200  & 100& 0.01 & $6912^3$ & $7.0\times 10^8$\\
       &       &       & 5  & 0.01 & $13824^3$& $3.0\times 10^5$\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure*}
\centering
  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
 \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}. The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$. The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel. The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
\label{fig.ccc}
\end{figure*}
\subsection{Power spectrum}
In S2048S and S2048L are two simulations with $2048^3$ particles on small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes. We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$. For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$. More generally we define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
\begin{equation}\label{eq.cps}
	\left\langle \delta_\alpha^\dagger({\bs k})\delta_\beta({\bs k}') \right\rangle=
    (2\pi)^3 P_{\alpha\beta}(k){\bs \delta}_{\rm 3D}({\bs k}-{\bs k}'),
\end{equation}
where ${\bs \delta}_{\rm 3D}$ is the three-dimensional Dirac delta function. In cosmology we usually consider the dimensionless power spectrum $\Delta^2_{\alpha\beta}(k)\equiv k^3 P_{\alpha\beta}(k)/(2\pi^2)$. The cross correlation coefficient is defined as
\begin{equation}\label{eq.ccc}
	\xi(k)\equiv P_{\alpha\beta}/\sqrt{P_{\alpha\alpha}P_{\beta\beta}}.
\end{equation}

@@

\section{Discussion and conclusion}\label{s.discussion}
We present an open-source parallel integer-based $N$-body algorithm that needs use very low memory usage, approaching 6 byte per particle (6 bpp). The accuracy of this 6pp configuration is acceptable on most scales of our interest. The accuracy is also adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms.

We analyze the deviation of power spectrum and de-correlation caused by integer-1 based algorithms in two simulations S2048S and S2048L. They have comparable number of particles per computing as the TianNu simulation. S2048S focuses on mass resolution -- it simulates a very small physical volume and its mass resolution $m_p=4\times 10^8\Msun$ is even higher than TianNu ($7\times 10^8\Msun$ for its CDM particles), whereas S2048L focuses on simulation volume and have higher particle mass. In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.

@@

\begin{figure}
\centering
  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
 \caption{\tcb{We don't need this figure in submission.} The solid lines show the power spectra of S2048L and S2048S at $z=0$, and dotted lines show the $z=1$ counterparts. For $z=0$, the two data tips are at $k=0.2\ k_{\rm Nyquist}$, where the power spectrum is suppressed by more than 20\% compared to higher resolution simulations.}
\label{fig.power}
\end{figure}

>>>>

As a $n_\chi=1$, 1D ($d=1$), 4-coarse-cell ($N_c=4$) example, if $$\chi_1=(-128,127,0,60),$$ particle number density $${\rho_c^{\rm 1D}}=(1,0,2,1),$$ then in unit of coarse cells, the accurate positions of these four particles are $$x_1=(0.001953125, 2.998046875, 2.501953125, 3.736328125).$$

@@

We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass. These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE}. Using different number of tiles per image gives exact same results. We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.

\begin{table}[]
\centering
\caption{Simulation configurations}
\label{t.sim}
\begin{tabular}{lrrrrrr}
\hline
& \multicolumn{6}{c}{Configurations}\\
\cline{2-7}
Name  & $N_{\rm node}$ & $L/({\rm Mpc}/h)$ & $z_i$ & $z_{\rm ckpt}$ & $N_p$ & $m_p/\Msun$ \\
\hline
S512   & 8     & 200   & 49 & 0    & $512^3$  & $7.5\times 10^9$    \\
S256   & 1     & 80    & 49 & 0    & $256^3$  & $3.8\times 10^9$    \\
S2048S & 64    & 400   & 49 & 1,0  & $2048^3$ & $9.4\times 10^8$    \\
S2048L & 64    & 1200  & 49 & 1,0  & $2048^3$ & $2.5\times 10^{10}$ \\
\hline
TianNu & 13824 & 1200  & 100& 0.01 & $6912^3$ & $6.9\times 10^8$\\
       &       &       & 5  & 0.01 & $13824^3$& $3.2\times 10^5$\\
TianZero & 13824 & 1200 & 100& 0.01 & $6912^3$ & $7.0\times 10^8$\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure*}
\centering
  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
 \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}. The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$. The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel. The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
\label{fig.ccc}
\end{figure*}
\subsection{Power spectrum}
In S2048S and S2048L are two simulations with $2048^3$ particles on small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes. We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$. The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}) to study halo properties. On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, reconstruction of baryonic acoustic oscillation (BAO) \citep{2017ApJ...841L..29W} etc.

For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$. More generally we define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
\begin{equation}\label{eq.cps}
	\left\langle \delta_\alpha^\dagger({\bs k})\delta_\beta({\bs k}') \right\rangle=
    (2\pi)^3 P_{\alpha\beta}(k){\bs \delta}_{\rm 3D}({\bs k}-{\bs k}'),
\end{equation}
where ${\bs \delta}_{\rm 3D}$ is the three-dimensional Dirac delta function. In cosmology we usually consider the dimensionless power spectrum $\Delta^2_{\alpha\beta}(k)\equiv k^3 P_{\alpha\beta}(k)/(2\pi^2)$. The cross correlation coefficient is defined as
\begin{equation}\label{eq.ccc}
	\xi(k)\equiv P_{\alpha\beta}/\sqrt{P_{\alpha\alpha}P_{\beta\beta}}.
\end{equation}

@@

\section{Discussion and conclusion}\label{s.discussion}
We present an parallel integer-based $N$-body algorithm. This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g. \cite{2017PhRvD..95d3501Y,2017ApJ...841L..29W,2017MNRAS.469.1968P}. It needs very low memory usage, approaching 6 byte per particle (6 bpp). The accuracy of this 6pp configuration is acceptable on most scales of our interest, e.g. BAO. The accuracy is also adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms.

We analyze the deviation of power spectrum and de-correlation caused by integer-1 based algorithms in two simulations S2048S and S2048L. They have comparable number of particles per computing as the TianNu simulation. S2048S focuses on mass resolution -- comparable to TianNu and TianZero simulations, whereas S2048L focuses on simulation volume and have higher particle mass. In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.

@@

\begin{figure}
\centering
  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
 \caption{\tcb{We don't need this figure in submission. However we need to compare the results with nonlinear power spectrum by {\tt HALOFIT}.} The solid lines show the power spectra of S2048L and S2048XS ($L=300\ {\rm Mpc}/h$) at $z=0$, and dotted lines show the $z=1$ counterparts. For $z=0$, the two data tips are at $k=0.2\ k_{\rm Nyquist}$, where the power spectrum is suppressed by more than 20\% compared to higher resolution simulations.}
\label{fig.power}
\end{figure}
